An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2022-01-28 paper.tex
abstract and intro

%% - \Abstract{} almost as usual
\Abstract{
  Likelihood-free inference (LFI) or Approximate Bayesian Computation (ABC) is the general framework for performing the inference in statistical models with intractable likelihood. In general, the intractability of the likelihood makes the inference problem quite challenging, therefore, many LFI methods either lack accuracy or are pretty inefficient. Robust Optimisation Monte Carlo (ROMC) is a highly-parallelizable approximate Bayesian computation (ABC) technique. In this paper, we present the implementation of ROMC in the \proglang{Python} package ELFI. Our implementation can be exploited in two directions. A scientist may use it as an out-of-the-box ABC algorithm for performing the inference in an LFI problem; we provide an easy-to-use API harmonized with the rest of the ELFI framework. Therefore, it is directly comparable with the rest of the methods included in ELFI. On the other direction, we have carefully designed ROMC in isolated algorithmic steps. Therefore, we allow the researcher to experiment with ROMC, altering or replacing parts of the algorithm with custom-created functionalities. In both scenarios, the algorithmic steps of ROMC can run in a fully-parallelizable manner, exploiting all CPU cores. We also provide some functionalities for visualization and evaluation needs. Finally, we test the robustness of our implementation on some typical LFI models that ship.

@@

Optimization Monte Carlo (OMC) proposed by [..] was a novel ABC approach approximating the posterior distribution. Afterward, Forneron and Ng provided a similar method under "reverse sampler ". The key idea is turning the stochastic data-generating mechanism into a set of fully-parallelizable deterministic optimization processes. Robust OMC (ROCM) provided an upgrade in OMC by locating and resolving some fundamental failure modes of the latter.

In this paper, we present an extendable and parallelizable implementation of ROMC at the Python package 'Engine for Likelihood-Free inference (ELFI) .' We have followed specific design principles for making the implementation extendible. Each step of the method is designed as a fully-isolated and parallelizable step; The user may easily replace parts of the technique with custom components without the backbone method collapse. Therefore, our implementation is a friendly tool for the research community to experiment with alterations of ROMC. Apart from the main functionalities, we also provide essential visualization and evaluation tools for gaining insights while fitting data. Finally, we verify the robustness of the implementation on the LFI examples supplied by the ELFI package.

>>>>

%% - \Abstract{} almost as usual
\Abstract{ Performing the inference in statistical models with
  intractable likelihood is challenging; most likelihood-free
  inference (LFI) methods encounter problems in accuracy and
  efficiency. In this paper, we present the implementation of Robust
  Optimisation Monte Carlo (ROMC) in the Python package ELFI. ROMC is
  a novel and efficient (highly-parallelizable) LFI technique that
  provides accurate weighted samples from the posterior.  Our
  implementation can be used in two ways. A scientist may use it as an
  out-of-the-box LFI algorithm. We provide an easy-to-use API
  harmonized with the rest of the ELFI framework; therefore, it is
  straightforward to compare it with all the methods included in
  ELFI. On the other direction, we have carefully split ROMC into
  isolated algorithmic steps. Therefore, we allow the researcher to
  experiment, replacing components of the algorithm with
  custom-designed subparts. In both scenarios, the algorithmic steps
  of ROMC run in a fully-parallelizable manner, exploiting all CPU
  cores. We also provide helpful functionalities for (i) inspecting
  the inference process and (ii) evaluating the approximated
  posterior. Finally, we test the robustness of our implementation on
  some typical LFI examples.

@@

Optimization Monte Carlo (OMC) proposed by~\citep{Meeds2015} was a
novel LFI approach for approximating the posterior
distribution. Afterwards, \citep{Forneron2016} provided a similar
method under `reverse sampler'. The key idea is turning the stochastic
data-generating mechanism into a set of fully-parallelizable
deterministic optimization processes. In their work,
~\citep{Ikonomov2019}, located some critical failure modes of OMC, and
proposed Robust OMC (ROMC) an alternative version of OMC with the
appropriate improvements.

In this paper, we present an extendible and parallelizable
implementation of ROMC at the \proglang{Python} package
\proglang{Engine for Likelihood-Free inference (ELFI)}. We follow
specific principles for making the implementation extendible. As we
will describe analytically in the next chapter, ROMC can be understood
not only as a specific algorithm, but mainly as a framework for
obtaining weighted samples from the posterior; it defines some
algorithmic task that must be fulfilled, without enforcing which
specific algorithm must be used for the tasks. Therefore, a researcher
may adopt ROMC's backbone approach, but experiment with alternative
methods to solve the distinct tasks. We have designed ROMC for
faciliating such alterations. Finally, we have tested the accuracy and
the efficienct of our implementation on the LFI examples supported by
the \proglang{ELFI} package.
